---
title: "Data analysis of developer responses"
output:
  html_notebook: default
  html_document: default
  pdf_document: default
---


# Introduction

# Load libraries

```{r,eval=TRUE,include=FALSE}
library(d3heatmap)
library(vcd)
library(GGally)
library(RColorBrewer)
library(graphics)
library(NbClust)
```

# Load data

```{r}
data <- read.csv( file = "./clus_data.csv", stringsAsFactors = FALSE )
clus_data <- read.csv("clus_data.csv")
clus_data <- data.frame(clus_data)
numericColumns <- sapply(data, is.numeric)
cols <- c(7:20, 26:35, 37:42) # Selecting the numerical columns
clus_data_selected <- clus_data[, cols]
str(clus_data_selected)
# and a scaled version of that with NA's removed
clus_data_scaled <- scale(na.omit(clus_data_selected))
```

```{r}
summary(as.data.frame(unclass(data)))
```

## Removing/marking outliers

The column JOBTIME has the values 9999 that would mess up some of the plots and statistics below.

```{r}
sum( data$JOBTIME > 200 , na.rm=T)
```

Turn values that too large into NA's.

```{r}
data$JOBTIME <- ifelse( data$JOBTIME > 200, NA, data$JOBTIME)
```

# Mosaic plots

It is a good idea to visualize the conditional probabilities between different columns.

## Role vs gender

```{r}
mosaic( ROLE ~ gender, data = data, color=2:3, rot_labels=c(45,45,45,45)  )
```
In the plot above we can see that the interviewed females are much more likely to be UX designers and managers than developers.

## Other trials

```{r}
mosaic( COMPANY ~ age_range, data = data, color=2:3, rot_labels=c(45,45,45,45)  )
```

## Role vs age

From the plots in this sub-section we can see:

- that there are no (interviewed) managers between 21-30;

- people older than 50 are much less likely to work in the industry (or answer interview questions);

- people older than 50 are much more likely to be managers, compared to the other age groups.

```{r}
mosaic( age_range ~ ROLE, data, direction = "v", rot_labels=c(45,45,45,45) )
```
```{r}
mosaic( ROLE ~ age_range, data, rot_labels=c(45,45,45,45) )
```



# Clustering

Using hierarchical clustering and visualizing with a heat-map.

```{r}
#inv <- clus_data_selected[, 15:20]
#class(inv)
d3heatmap( clus_data_selected, Rowv = TRUE, Colv = FALSE, scale = "column", na.rm = TRUE, dendrogram = "row", k_row = 4)
```

### Correlations benween coliumns 

Correlations for all numerical columns:

```{r}
qData <- data[, grepl("INV",colnames(data)) & numericColumns ]
qData <- data[, numericColumns ]
#qData <- qData[ rowSums( is.na(qData) ) < 6, ]
GGally::ggcorr( qData, method = c("pairwise","spearman") )
```

It is a good idea to explain the correlations that are close to 1 and -1. Do they make sense, etc.

Correlations for selected numerical columns:

```{r}
# Numerical columns starting with "INV".
qData <- data[, grepl("^INV",colnames(data)) & numericColumns ]
# Removing rows that have too many NA's.
qData <- qData[ rowSums( is.na(qData) ) < 6, ]
GGally::ggcorr( qData, method = c("pairwise","spearman") )
```
## try http:// if https:// URLs are not supported
source("https://bioconductor.org/biocLite.R")
biocLite("ComplexHeatmap")
