---
title: "Data analysis of developer responses"
output:
  html_document: default
  html_notebook: default
---


# Introduction

# Load libraries

```{r,eval=TRUE,include=FALSE}
library(d3heatmap)
library(vcd)
library(GGally)
library(dplyr)
library(reshape2)
library(ggplot2)
library(RColorBrewer)
```

# Load data

```{r}
data <- read.csv( file = "./clus_data.csv", stringsAsFactors = FALSE )
#write.csv( x = data[,-1], file = "./clus_data_new.csv", row.names = FALSE )

numericColumns <- sapply(data, is.numeric)
```

```{r}
summary(as.data.frame(unclass(data)))
```

## Removing/marking outliers

The column JOBTIME has the values 9999 that would mess up some of the plots and statistics below.

```{r}
sum( data$JOBTIME > 200 , na.rm=T)
```

Turn values that too large into NA's.

```{r}
data$JOBTIME <- ifelse( data$JOBTIME > 200, NA, data$JOBTIME)
```

# Mosaic plots

It is a good idea to visualize the conditional probabilities between different columns.

## Role vs gender

```{r}
data1 <- data
data1 <- data.frame(data1)
data1[data1=="Ericsson"]<-"Company A"
data1[data1=="F-secure"]<-"Company B"
data1[data1=="Vaadin"]<-"Company C"
data1[data1=="Reaktor"]<-"Company D"

# data2 <- subset(data1, condexp %in% c("Never", "Rarely", "Occasionally", "Yes, actively"))
data1$condexp = factor(data1$condexp, levels=c("Never", "Rarely", "Occasionally", "Yes, actively"))
data2 <- subset(data1, ROLE %in% c("Developer", "Manager", "UX designer"))


mosaic( condexp ~ ROLE, data = data2, rot_labels=c(33,0,0,0), highlighting_fill = subs_pal[3:6], labeling_args = list(set_varnames = c(condexp=" ", ROLE=""), just_labels = c("left", 
                                                 "center", 
                                                 "center", 
                                                 "right"), offset_labels = c(-0.72, 0, 0, 0))
      )
```
In the plot above we can see that the interviewed females are much more likely to be UX designers and managers than developers.

## Role vs age

From the plots in this sub-section we can see:

- that there are no (interviewed) managers between 21-30;

- people older than 50 are much less likely to work in the industry (or answer interview questions);

- people older than 50 are much more likely to be managers, compared to the other age groups.

```{r}
data1 <- data
data1 <- data.frame(data1)
data1[data1=="Ericsson"]<-"A"
data1[data1=="F-secure"]<-"B"
data1[data1=="Vaadin"]<-"C"
data1[data1=="Reaktor"]<-"D"

data2 <- subset(data1, ROLE %in% c("Developer", "Manager", "UX designer"))

#set_labels = list(ROLE = c(" ", " ", ""))
subs_pal <- colorspace::diverge_hcl(7)
#mosaic( ROLE ~ COMPANY, data = data2, highlighting_fill = subs_pal[1:4],  rot_labels=c(0,0,45,45), labeling_args = list(set_varnames = c(ROLE=" ", COMPANY=""))  )

mosaic( ROLE ~ COMPANY, data2,  gp = gpar(fill = c("lightsteelblue2" , "royalblue3", "snow4") ), direction = "v", rot_labels=c(0,0,45,45), labeling_args = list(set_varnames = c(ROLE=" ", COMPANY=""))) 
```
```{r}
mosaic( ROLE ~ age_range, data, rot_labels=c(45,45,45,45) )
```



# Clustering

Using hierarchical clustering and visualizing with a heat-map.

```{r}
d3heatmap( data[,numericColumns][,-1], Rowv = TRUE, Colv = FALSE, scale = "column", na.rm = TRUE )
```

### Correlations benween coliumns 

Correlations for all numerical columns:

```{r}
qData <- data[, grepl("INV",colnames(data)) & numericColumns ]
qData <- data[, numericColumns ]
#qData <- qData[ rowSums( is.na(qData) ) < 6, ]
GGally::ggcorr( qData, method = c("pairwise","spearman") )
```

It is a good idea to explain the correlations that are close to 1 and -1. Do they make sense, etc.

Correlations for selected numerical columns:

```{r}
# Numerical columns starting with "INV".
qData <- data[, grepl("^INV",colnames(data)) & numericColumns ]
# Removing rows that have too many NA's.
qData <- qData[ rowSums( is.na(qData) ) < 6, ]
GGally::ggcorr( qData, method = c("pairwise","spearman") )
```

# Average interviewees per class


## Average interviewee per company

```{r}
x <- c("INVA", "INVB", "INVC", "INVD", "INVE",  "INVF",  "INVG")
#"INVH", "INVI", "INVJ", "INVK", "INVL", "INVM", "INVN"
#"INVA", "INVB", "INVC", "INVD", "INVE",  "INVF",  "INVG"
qColNames <- colnames(data)[ grepl(paste(x, collapse = "|"), colnames(data)) & numericColumns ]
#qColNames <- colnames(data)[ grepl("^USER", colnames(data)) & numericColumns ]
colNamesToColIndsRules <- setNames( 1:length(qColNames), qColNames )
qDF <- 
  melt( data = data, id.vars = c("COMPANY") ) %>% 
  dplyr::filter( variable %in% qColNames ) %>%
  dplyr::filter(!is.na(value)) %>%
  dplyr::filter(!is.na(COMPANY)) 
#%>%
  #dplyr::filter(!is.na(ROLE))

group_means <- qDF %>%
  group_by(COMPANY, variable) %>%
  dplyr::summarise(mean = mean(as.numeric(value)))

group_means[group_means=="Ericsson"]<-"A"
group_means[group_means=="F-secure"]<-"B"
group_means[group_means=="Vaadin"]<-"C"
group_means[group_means=="Reaktor"]<-"D"


#levels(group_means$variable)
group_means$variable <- factor(group_means$variable)
#Reaktor <- subset(group_means, COMPANY=="A")
#ggplot(subset(Reaktor, ROLE %in% c("Developer", "Manager", "UX designer")),
group_means <- data.frame(group_means)
#subset(group_means, COMPANY %in% c("Developer", "Manager", "UX designer"))
ggplot(group_means, aes(x=as.numeric(variable), y=mean, color=COMPANY, linetype=COMPANY)) + geom_line(size=1) +
  scale_linetype_manual(values=c("solid", "twodash", "dotted", "longdash")) +
  scale_x_continuous(labels=as.character(group_means$variable), breaks=as.numeric(group_means$variable))+
  theme(axis.text=element_text(size=12))+
  theme(axis.title.x=element_blank(),
        #axis.text.y=element_blank(),
        #axis.ticks.x=element_blank(),
        axis.title.y=element_blank())  + coord_cartesian(ylim=c(1, 5)) +
  theme(legend.key.size = unit(3,"line"),  legend.text=element_text(size=12))

  

```

## Average interviewee per company

```{r}
#x <- c("INVA", "INVC", "INVD", "INVE", "INVF",  "INVN",  "INVL")
#qColNames <- colnames(data)[ grepl(paste(x, collapse = "|"), colnames(data)) & numericColumns ]
qColNames <- colnames(data)[ grepl("^X3.3", colnames(data)) & numericColumns ]
colNamesToColIndsRules <- setNames( 1:length(qColNames), qColNames )
qDF <- 
  melt( data = data, id.vars = c("ROLE") ) %>% 
  dplyr::filter( variable %in% qColNames ) %>%
  dplyr::filter(!is.na(value)) %>%
  dplyr::filter(!is.na(ROLE)) 
#%>%
  #dplyr::filter(!is.na(ROLE))

group_means <- qDF %>%
  group_by(ROLE, variable) %>%
  dplyr::summarise(mean = mean(as.numeric(value)))

group_means[group_means=="Ericsson"]<-"A"
group_means[group_means=="F-secure"]<-"B"
group_means[group_means=="Vaadin"]<-"C"
group_means[group_means=="Reaktor"]<-"D"


#levels(group_means$variable)
group_means$variable <- factor(group_means$variable)
#Reaktor <- subset(group_means, COMPANY=="A")
#ggplot(subset(Reaktor, ROLE %in% c("Developer", "Manager", "UX designer")),
group_means <- data.frame(group_means)
#subset(group_means, COMPANY %in% c("Developer", "Manager", "UX designer"))
ggplot(subset(group_means, ROLE %in% c("Developer", "Manager", "UX designer")), aes(x=as.numeric(variable), y=mean, color=ROLE, linetype=ROLE)) + geom_line(size=1) +
  scale_linetype_manual(values=c("solid", "twodash", "dotted", "longdash")) +
  scale_x_continuous(labels=as.character(group_means$variable), breaks=as.numeric(group_means$variable))+
  theme(axis.text=element_text(size=12))+
  theme(axis.title.x=element_blank(),
        #axis.text.y=element_blank(),
        #axis.ticks.x=element_blank(),
        axis.title.y=element_blank()) +
  coord_cartesian(ylim=c(1, 5)) +
  theme(legend.key.size = unit(3,"line"),  legend.text=element_text(size=12))

  

```

## Average interviewee per role

```{r}
qDF <- 
  melt( data = data, id.vars = c("ROLE") ) %>% 
  dplyr::filter( variable %in% qColNames ) %>%
  dplyr::filter(!is.na(value)) %>%
  dplyr::filter(!is.na(ROLE)) 
  
group_means <- qDF %>%
  group_by(ROLE, variable) %>%
  dplyr::summarise(mean = mean(as.numeric(value)))

group_means$variable <- factor(group_means$variable)

ggplot(subset(group_means, ROLE %in% c("Developer", "Manager", "UX designer")), aes(x=as.numeric(variable), y=mean, color=ROLE, linetype=ROLE)) + geom_line(size=1) + scale_linetype_manual(values=c("solid", "twodash", "dotted")) +
    theme(axis.text=element_text(size=12)) +
  scale_x_continuous(labels=as.character(group_means$variable), breaks=as.numeric(group_means$variable)) +
  theme(axis.title.x=element_blank(),
        #axis.text.y=element_blank(),
        #axis.ticks.x=element_blank(),
        axis.title.y=element_blank()) +
   scale_y_discrete(limits = c(1, 4)) + scale_y_discrete(limits= c(1, 2, 3, 4)) +
  theme(legend.key.size = unit(3,"line"),  legend.text=element_text(size=12))
``` 
